---
title: "BACSAL_new"
author: "migs"
date: "2023_2_6"
output: html_document
---
## Loading packages
```{r, echo = T, message=FALSE}
library(ggplot2)
library(magrittr)
library(tidyverse)
```


## Importing BACSAL data set
```{r, echo=T, results = 'hide'}
library(readr)
bsdata <- read_csv("complete_BACSAL_data.csv", show_col_types = F)
head(bsdata)
```

## Renaming long column names
```{r, echo=T, results = 'hide'}
bsdata <- bsdata %>% 
  rename(L01 = `0<.L.<=0.5000000`,
         L02 = `0.5000000<.L.<=1.0000000`,
         L03 = `1.0000000<.L.<=1.5000000`,
         L04 = `1.5000000<.L.<=2.0000000`,
         L05 = `1.5000000<.L.<=2.0000000`,
         L06 = `2.0000000<.L.<=2.5000000`,
         L07 = `2.5000000<.L.<=3.0000000`,
         L08 = `3.0000000<.L.<=3.5000000`,
         L09 = `3.5000000<.L.<=4.0000000`,
         L10 = `4.0000000<.L.<=4.5000000`,
         L11 = `.L.>4.5000000`)
```



## Creating & converting variables

Root-shoot ratio (RTS) - the ratio of root dry biomass over shoot dry biomass)
Note: Mou et al. 2013 uses rts, has no units (dimensionless)
```{r, echo = T, results='hide'}
bsdata %>% 
  select(`Root Mass (g)`, `Shoot mass (g)`) %>% 
  mutate(rts = `Root Mass (g)` / `Shoot mass (g)`)
```
Specific Root Length (SRL) - root length divided by root dry mass
Note: Funk et al. 2016 uses units (cm / mg), while Kramer-Walter et al. 2016 uses units (m/g). Will use KW units.
```{r, echo = T, results='hide'}
bsdata %>% 
  select(Code, `Length(cm)`, `Root Mass (g)`) %>% 
  mutate(srl = (`Length(cm)`/100)/ `Root Mass (g)`)
```
Root Tissue Density (RTD) - root mass/root volume
Note: Kramer-Walter et al. 2016 uses mg/mm^3

```{r, echo = T, results='hide'}
bsdata %>% 
  select(Code, `Root Mass (g)`, `RootVolume(cm3)`) %>% 
  mutate(rtd = ((`Root Mass (g)`*1000)/(`RootVolume(cm3)`*1000)))
```

Specific Root Area (SRA) -root area divided by dry root mass
Note: Sun et al. 2020 uses units (cm^2 / g)
```{r, echo = T, results='hide'}
bsdata %>% 
  select(Code, `SurfArea(cm2)`, `Root Mass (g)`) %>% 
  mutate(sra = `SurfArea(cm2)` / `Root Mass (g)`)
```
<!-- Relative Growth Rate, Aboveground (RGR) aka volumetric root data (delta: cm^3) -->
<!-- ```{r, echo = T, results='hide'} -->
<!--  bsdata %>%  -->
<!--    select(Code, f.height, f.width1, f.width2, i.height, i.width1, i.width2) %>%  -->
<!--    mutate(volrgr = (((f.height/10) * (f.width1/10) * (f.width2/10)) - ((i.height/10) * (i.width1/10) * (i.width2/10)))) -->
<!-- ``` -->

RGR, according to Kramer-Walter et al. 2016 (t = 35 days)
```{r, echo = T, results='hide'}
bsdata %>% 
  select(Code, i.height, f.height) %>% 
  mutate(kwrgr = ((log(f.height)) - (log(i.height)))/35)
```


UNFINISHED WORK BELOW
Fine:Coarse Root Ratio -need to do some digging in McCormack 2015
Note: McCormack et al. 2015 uses arbitrary 2 mm as cutoff for fine:coarse roots ratio and questions this cutoff. Blume-Werry et al. 2017 uses log transformed fine/coarse root ratio (dimensionless).
Update Note: Maybe showing root histograms? ie, length? would be logarithmic, orders of magnitude are too greatly varied between <.5 mm and >4.5 mm

Creating histogram: 
Xaxis - Root diam. classification (from <0.5 mm to >4.5 mm)
Yaxis - root length total (log transformed?)
Group by sites (biggest source of variance), take means, and add add error bars
# ```{r, echo = T, results='hide'}
# #Reorganizing root data into one data frame
# # bsdata <- structure(list(diamclass = c('0<.L.<=0.5000000', '0.5000000<.L.<=1.0000000')), .Names = c('diamclass', 'rlength'), class = 'data.frame')
# 
# bsdata %>% 
#   select(Code, Origin, Treatment, 
#         '0<.L.<=0.5000000', 
#         '0.5000000<.L.<=1.0000000',
#         `RootVolume(cm3)`,
#         `Root Mass (g)`
#         ) %>%
#   ggplot(aes(x = log(`Root Mass (g)`), color = Origin)) +
#     geom_density()
# 
# ```

Root-shoot ratio, by Nitrogen
```{r, echo = T, results='show'}
bsdata %>% 
  select(Code, Origin, Treatment, Water, Nitrogen, `Root Mass (g)`, `Shoot mass (g)`) %>% 
  mutate(rts = `Root Mass (g)` / `Shoot mass (g)`) %>% 
  ggplot() +
    aes(x = Origin, y = rts, fill = Nitrogen) +
    geom_boxplot() +
    ggtitle("Root-shoot ratio, grouped by Nitrogen") +
    labs(y = "Root-to-Shoot")
```

Root-shoot ratio, by Origin
```{r, echo = T, results='show'}
bsdata %>% 
  select(Origin, `Root Mass (g)`, `Shoot mass (g)`) %>% 
  mutate(rts = `Root Mass (g)` / `Shoot mass (g)`) %>% 
  ggplot() +
    aes(x = Origin, y = rts, fill = Origin) +
    geom_boxplot() +
    ggtitle("Root-shoot ratio, grouped by Origin") +
    labs(y = "Root-to-Shoot")
```

According to Funk et al. 2017, a soft trait like SLA "can explain a large portion of the variation in RGR across a large range of herbaceous and woody plant species". can SRA (specific root area) do the same for root volume?
Relative Growth Rate, Aboveground (RGR) aka volumetric root data (delta: cm^3)
```{r, echo = T, results='show'}
#Root volume vs. specific root area (sra)
bsdata %>% 
  select(Code, Origin, `SurfArea(cm2)`, Origin, `Root Mass (g)`, `RootVolume(cm3)`) %>% 
  mutate(sra = `SurfArea(cm2)` / `Root Mass (g)`) %>% 
  ggplot() +
    aes(x = sra, y = `RootVolume(cm3)`, color = Origin) +
    geom_point(na.rm = T) +
    scale_x_continuous(limits = c(0,1800)) +
    scale_y_continuous(limits = c(0,36))
```
```{r, echo = T, results='show'}
#RGR vs. root diameter
bsdata %>% 
  select(Code, `AvgDiam(mm)`, Origin, f.height, f.width1, f.width2, i.height, i.width1, i.width2) %>%
  mutate(logvolrgr = log((((f.height/10) * (f.width1/10) * (f.width2/10)) - ((i.height/10) * (i.width1/10) * (i.width2/10))))) %>% 
  ggplot() +
    aes(x = `AvgDiam(mm)`, y = logvolrgr, color = Origin) +
    geom_jitter(na.rm = T) +
    scale_x_continuous(limits = c(0.3, 1.3))
```
Kramer-Walter et al. 2016 uses bivariate comparisons between RGR and various root traits
```{r, echo = T, results='show'}
#KWrgr vs. AvgDiam
bsdata %>% 
  select(Code, `AvgDiam(mm)`, Origin, f.height, f.width1, f.width2, i.height, i.width1, i.width2) %>%
 mutate(kwrgr = ((log(f.height)) - (log(i.height)))/35) %>% 
  ggplot() +
    aes(x = `AvgDiam(mm)`, y = kwrgr, color = Origin) +
    geom_jitter(na.rm = T)
```


```{r, echo = T, results='show'}
#KWRGR vs. SRL
bsdata %>% 
  select(Code, `Length(cm)`, `Root Mass (g)`, Origin, f.height, f.width1, f.width2, i.height, i.width1, i.width2) %>%
  mutate(kwrgr = ((log(f.height)) - (log(i.height)))/35) %>% 
  mutate(srl = (`Length(cm)`/100)/ `Root Mass (g)` ) %>% 
  ggplot() +
    aes(x = srl, y = kwrgr, color = Origin) +
    geom_point(na.rm = T) +
    scale_x_continuous(limits = c(0,100))
```


```{r, echo = T, results='show'}
#KWrgr vs. RTD
bsdata %>% 
  select(Code, `Root Mass (g)`, `RootVolume(cm3)`, `Root Mass (g)`, Origin, f.height, i.height) %>% 
  mutate(rtd = ((`Root Mass (g)`*1000)/(`RootVolume(cm3)`*1000))) %>%
  mutate(kwrgr = ((log(f.height)) - (log(i.height)))/35) %>% 
  ggplot() +
    aes(x = rtd, y = kwrgr, color = Origin) +
    geom_point(na.rm = T)
```

<!-- #After D-Lab consultation, 2/9/23 -->
<!-- ```{r} -->
<!-- bsdata %>%  -->
<!--   select(Code, Origin, `0<.L.<=0.5000000`, ) -->


<!-- ``` -->





```{r}
bsdatacat <- bsdata %>% 
  pivot_longer(L01:L11, names_to = 'Lcategory', values_to = 'LengthCat');

#Non-normalized
bsdatacat %>%
  # select(Code, Origin, LengthCat, Lcategory) %>% 
    ggplot() +
    aes(x = Lcategory, y = 1+log(LengthCat), fill = Origin) +
    geom_boxplot()


#Normalized by Total Length
bsdatacat %>%
  select(Code, Origin, `Length(cm)`, LengthCat, Lcategory) %>% 
  mutate(NormaLCat = LengthCat /`Length(cm)`) %>% 
    ggplot() +
    aes(x = Lcategory, y = (NormaLCat), fill = Origin) +
    geom_boxplot()
#Note: graphic looks weird. Can NormaLCat be transformed by something other than log?


# Box-Cox plot might be helpful!
```















