---
title: "BACSAL_new"
author: "migs"
date: "2022-10-31"
output: html_document
---
## Loading packages
```{r, echo = T, message=FALSE}
library(ggplot2)
library(magrittr)
library(tidyverse)
```



## Importing BACSAL data set
```{r, echo=T, results = 'hide'}
library(readr)
bsdata <- read_csv("complete_BACSAL_data.csv", show_col_types = F)
head(bsdata)
```

## Creating & converting variables

Root-shoot ratio (RTS) - the ratio of root dry biomass over shoot dry biomass)
Note: Mou et al. 2013 uses rts, has no units (dimensionless)
```{r, echo = T, results='hide'}
bsdata %>% 
  select(Code, `Root Mass (g)`, `Shoot mass (g)`) %>% 
  mutate(rts = `Root Mass (g)` / `Shoot mass (g)`)

```

Specific Root Length (SRL) - root length divided by root dry mass
Note: Funk et al. 2016 uses units (cm / mg), while Kramer-Walter et al. 2016 uses units (m/g). Will use KW units.
```{r, echo = T, results='hide'}
bsdata %>% 
  select(Code, `Length(cm)`, `Root Mass (g)`) %>% 
  mutate(srl = (`Length(cm)`/100)/ `Root Mass (g)` )

```

Specific Root Area (SRA) -root area divided by dry root mass
Note: Sun et al. 2020 uses units (cm^2 / g)
```{r, echo = T, results='hide'}
bsdata %>% 
  select(Code, `SurfArea(cm2)`, `Root Mass (g)`) %>% 
  mutate(sra = `SurfArea(cm2)` / `Root Mass (g)`)

```

Fine:Coarse Root Ratio -need to do some digging in McCormack 2015
Note: McCormack et al. 2015 uses arbitrary 2 mm as cutoff for fine:coarse roots ratio and questions this cutoff. Blume-Werry et al. 2017 uses log transformed fine/coarse root ratio (dimensionless)
```{r, echo = T, results='hide'}
bsdata %>% 
  select(Code, '0<.L.<=0.5000000')

```












