---
title: "BACSAL_new"
author: "migs"
date: "2023_2_6"
output: html_document
---
## Loading packages
```{r Package Library, echo=T, message=FALSE}
library(readr)
library(ggplot2)
library(magrittr)
library(tidyverse)
library(ggsci)
library(factoextra)
library(lme4)
library(lmerTest)
library(htmltools)
library(devtools)
library(caret)
library(mice)
library(sciplot)
library(psych)
library(agricolae)
library(Plasticity)
```

## Importing BACSAL data set
```{r Data Import, echo=T, results = 'hide'}
bsdata <- read_csv("complete_BACSAL_data.csv", show_col_types = F)
# view(bsdata)
```


## Renaming long column names
```{r Column Renaming, echo=T, results = 'hide'}
bsdata <- bsdata %>% 
  rename(L01 = "0<.L.<=0.5000000",
         L02 = "0.5000000<.L.<=1.0000000",
         L03 = "1.0000000<.L.<=1.5000000",
         L04 = "1.5000000<.L.<=2.0000000",
         L05 = "2.0000000<.L.<=2.5000000",
         L06 = "2.5000000<.L.<=3.0000000",
         L07 = "3.0000000<.L.<=3.5000000",
         L08 = "3.5000000<.L.<=4.0000000",
         L09 = "4.0000000<.L.<=4.5000000",
         L10 = ".L.>4.5000000")
```


## Creating & converting variables
```{r Calculating Trait Values, echo = T, results='hide'}
#Add in all created traits, remove all missing values
bsdata_complete <- bsdata %>% 
  mutate(rts = RootMassg / ShootMassg) %>% #Root-to-Shoot
  mutate(srl = (Lengthcm/100) / RootMassg) %>% #SRL
  mutate(sra = SurfAreacm2 / RootMassg) %>% #SRA
  mutate(rtd = ((RootMassg*1000)/(RootVolcm3*1000))) %>% #RTD
  mutate(kwrgr = ((log(fHeight)) - (log(iHeight)))/35) #RGR
# head(bsdata_complete)
```
```{r Recategorization, echo = T, results = 'hide'}
bsdata_complete$Origin <- as.factor(bsdata_complete$Origin)
bsdata_complete$Sex <- as.factor(bsdata_complete$Sex)
bsdata_complete$Nitrogen <- as.factor(bsdata_complete$Nitrogen)
bsdata_complete$Water <- as.factor(bsdata_complete$Water)
bsdata_complete$Parent <- as.factor(bsdata_complete$Parent)
bsdata_complete$Treatment <- as.factor(bsdata_complete$Treatment)
# str(bsdata_complete)

```





Root-shoot ratio (RTS) - the ratio of root dry biomass over shoot dry biomass)
Note: Mou et al. 2013 uses rts, has no units (dimensionless)
```{r, echo = T, results='hide'}
bsdata %>% 
  select(RootMassg, ShootMassg) %>% 
  mutate(rts = RootMassg / ShootMassg)
```
Specific Root Length (SRL) - root length divided by root dry mass
Note: Will use KW 2016 units (m/g).
```{r, echo = T, results='hide'}
bsdata %>% 
  select(Code, Lengthcm, RootMassg) %>% 
  mutate(srl = (Lengthcm/100) / RootMassg)
```
Root Tissue Density (RTD) - root mass/root volume
Note: Kramer-Walter et al. 2016 uses mg/mm^3
```{r, echo = T, results='hide'}
bsdata %>% 
  select(Code, RootMassg, RootVolcm3) %>% 
  mutate(rtd = ((RootMassg*1000)/(RootVolcm3*1000)))
```

Specific Root Area (SRA) -root area divided by dry root mass
Note: Sun et al. 2020 uses units (cm^2/g)
```{r, echo = T, results='hide'}
bsdata %>% 
  select(Code, SurfAreacm2, RootMassg) %>% 
  mutate(sra = SurfAreacm2 / RootMassg)
```

KWRGR - (relative height growth rate)
Note: Kramer-Walter et al. 2016, uses (ln(f)-ln(i))/(t) (BACSAL t = 35 days)
```{r, echo = T, results='hide'}
bsdata %>%
  select(Code, iHeight, fHeight) %>% 
  mutate(kwrgr = ((log(fHeight)) - (log(iHeight)))/35)
```

UNFINISHED WORK BELOW

Root-shoot ratio, by Nitrogen
Note: Add stat significant differences between Low & High with ***
Will use independet t-test (unpaired) for now - http://www.sthda.com/english/wiki/unpaired-two-samples-t-test-in-r
```{r Root-to-shoot, echo=T, results='show'}
#plotting rts against site aridity, separated by Nitrogen
bsdata_complete %>% 
  ggplot() +
  aes(x = AIKoppen, y = rts, outline = Origin, fill = Nitrogen) +
    geom_boxplot(na.rm = T ) + #removes missing values
    ggtitle("Root-shoot ratio, grouped by Nitrogen") +
    labs(y = "Root-to-Shoot Ratio", x = "Koppen Aridity Index") +
    scale_fill_discrete(labels = c("High N", "Low N"))

##Trying to subset each Origin to perform Welch T Test
#BOYD
boyd_rts <- bsdata %>% 
  mutate(rts = RootMassg / ShootMassg) %>% 
  subset(Origin == "BOYD")
t.test(data = boyd_rts, rts ~ Nitrogen, paired = F)
#BOYD: p<0.05, p-value = 2.531e-06, ***

#ELFR
elfr_rts <- bsdata %>% 
  mutate(rts = RootMassg / ShootMassg) %>% 
  subset(Origin == "ELFR")
t.test(data = elfr_rts, rts ~ Nitrogen, paired = F)
#ELFR: p<0.05, p-value = 5.808e-07, ***

#COAL
coal_rts <- bsdata %>% 
  mutate(rts = RootMassg / ShootMassg) %>% 
  subset(Origin == "COAL")
t.test(data = coal_rts, rts ~ Nitrogen, paired = F)
#COAL: p<0.05, p-value = 0.002649, ***
#All ORIGIN groups are significantly different between their high and low N groups


#Mixed-effects model:
#1 build hierarchical model (having isSingular warning when (1|Parent:Origin) random effect added)
#ask Greg how to deal w/ overfitting
#can keep water p-values, it'll avg across them, or can remove Water
hmod1 <- lmer(rts ~ Nitrogen*Water + AIKoppen + (1|Sex) + (1|Origin) , data = bsdata_complete)
hmod2 <- lmer(rts ~ Nitrogen*Water + AIKoppen + (1|Sex) + (1|Parent:Origin) , data = bsdata_complete)


#checking for differences in model
anova(hmod1, hmod2)

library(emmeans)
anova(hmod1)

#Pairwise comparisons
emmeans(hmod1, list(pairwise ~ Nitrogen), adjust = "tukey")


#plotting N-Water interaction effects
library(effects)
library(sjPlot)
eff.p1 <- effect("Nitrogen*Water", hmod1, KR=T)
plot(eff.p1)
plot(hmod1) #plotting residuals; output is NOT very random
#look at non-normal residuals; QQ-plot
qqnorm(resid(hmod1)) #pretty good! one outlier (check on BOYD-13, run with AND without)-if they aren't too different, it's ok! try removing it and seeing

```





Building a linear regression model for root-to-shoot ratio
```{r RTS Model Selection, echo=T, results='show'}
#Distribution Check
bsdata_complete %>% 
  ggplot() +
    aes(rts) +
    geom_histogram(na.rm = T,
                   aes(y = after_stat(density)), fill = "orange",
                   bins = 10) +
    geom_density(na.rm = T)
#Not very normal - has heavy outliers (from BOYD Low N group)
#what matters is residuals are normally dist, NOT the data itself

#Using lmer package, allows for mixed effects model
mod_rts <- lmer(rts ~ Water*Nitrogen + AIKoppen + (1 | Origin) + (1 | Parent:Origin), data = bsdata_complete)
#Origin and Parent:Origin are random effects, not what we're looking at
#Note: Adding in 'Sex' Variable produces a singular fit error, may be overleveraged (fewer degrees of freedom)

summary(mod_rts)
anova(mod_rts)


step_result <- step(mod_rts)

final_model <- get_model(step_result)
final_model
summary(final_model)
#formula = rts ~ Nitrogen + AIKoppen
plot(mod_rts)
plot(cooks.distance(mod_rts))
summary(final_model)$r.squared
```


Root-shoot ratio, by AIKoppen
```{r, echo = T, results='show'}
bsdata_complete %>% 
  ggplot() +
    aes(x = AIKoppen, y = rts, fill = Origin) +
    geom_boxplot(na.rm = T) + #removes missing values
    ggtitle("Root-shoot ratio") +
    labs(y = "Root-to-Shoot Ratio",
         x = "Koppen Aridity Index") +
    scale_fill_discrete(labels = c("BOYD", "ELFR", "COAL"))

#Create temp df for stats tests
origin_rts <- bsdata_complete$rts

#One-Way ANOVA to determine statistical significance + Tukey multiple pairwise-comparison
# rts.aov <- aov(rts ~ Origin, data = origin_rts)
# summary(rts.aov)
# TukeyHSD(rts.aov)
#Note: From Tukey test, ELFR & COAL are NOT statistically significant from each other
```


Fine:Coarse Root Ratio -need to do some digging in McCormack 2015
Note: McCormack et al. 2015 uses arbitrary 2 mm as cutoff for fine:coarse roots ratio and questions this cutoff. Blume-Werry et al. 2017 uses log transformed fine/coarse root ratio (dimensionless).
```{r Fine-Coarse Root Ratio,  echo = T, results='show'}
bsdata_complete %>% 
  mutate(froots = (L01 + L02 + L03 + L04)) %>% 
  mutate(croots = (L05 + L06 + L07 + L08 + L09 + L10)) %>% 
  mutate(fcr = froots/croots) %>% 
  ggplot() +
    aes(x = AIKoppen, y = log(fcr), fill = Origin, color = Nitrogen) +
    geom_boxplot(na.rm = T) +
      ggtitle("Fine-Coarse Root Ratio") +
        labs(y = "Fine-Coarse Root Ratio (f<=2mm<c)",
         x = "Koppen Aridity Index")


```


According to Funk et al. 2017, a soft trait like SLA "can explain a large portion of the variation in RGR across a large range of herbaceous and woody plant species". can SRA (specific root area) do the same for root volume?
```{r, echo = T, results='show'}
#Specific Root Area by AIKoppen
bsdata_complete %>% 
  ggplot() +
    aes(x = AIKoppen, y = sra, fill = Origin, color= Nitrogen) +
    geom_boxplot(na.rm = T) +
    scale_y_continuous(limits = c(0,1500)) +
    labs(y = "Specific Root Area (SRA) ",
         x = "Koppen Aridity Index")

#Create temp df for stats tests
origin_sra <- bsdata %>% 
  mutate(sra = SurfAreacm2 / RootMassg)

#One-Way ANOVA to determine statistical significance + Tukey multiple pairwise-comparison
sra.aov <- aov(sra ~ Origin, data = origin_sra)
summary(sra.aov)
TukeyHSD(sra.aov)

#Note: None of the SRA's are significantly different, ELFR & COAL are very aligned
```
SRA vs. KWRGR
```{r}
#Specific Root Area vs KW RGR
bsdata_complete %>% 
  ggplot() +
    aes(x = sra, y = kwrgr, color = Origin) +
    geom_point(na.rm = T) +
    scale_x_continuous(limits = c(0,1500)) +
    stat_ellipse(na.rm = T, 
                 lwd = 1.2) +
    labs(y = "Relative Growth Rate",
         x = "Specific Root Area (SRA)")
#Meaning: 3 sites have similar SRA distribution, but BOYD have slower growth rate based on similar SRA, COAL has highest
#run stats on this (BOYD has some slope)
```


```{r, echo = T, results='show'}
#RGR vs. root diameter
bsdata_complete %>% 
  ggplot() +
    aes(x = AvgDiammm, y = kwrgr, color = Origin) +
    geom_jitter(na.rm = T) +
    scale_x_continuous(limits = c(0, 1.3)) +
    stat_ellipse(na.rm = T, 
                 lwd = 1.2) +
    labs(y = "Relative Growth Rate",
         x = "Root Average Diameter (mm)")
```

```{r, echo = T, results='show'}
#KWRGR vs. Specific Root Length
bsdata_complete %>% 
  ggplot() +
    aes(x = srl, y = kwrgr, color = Origin) +
    geom_point(na.rm = T) +
    scale_x_continuous(limits = c(0,100)) +
    stat_ellipse(na.rm = T, 
                 lwd = 1.2) +
    labs(y = "Relative Growth Rate",
         x = "Specific Root Length (m/g)")
#High SRL signals more 'exploratory' growth patterns. In BOYD, a high SRL, more exploratory strategy is associated with higher RGR (performance), in ELFR, a low SRL more "conservative" root development is associated with higher RGR/performance. COAL has no clear relationship. 
```


```{r, echo = T, results='show'}
#KWrgr vs. RTD
bsdata_complete %>% 
  ggplot() +
    aes(x = rtd, y = kwrgr, color = Origin, shape = Nitrogen) +
    geom_point(na.rm = T) +
    stat_ellipse(na.rm = T, 
                 lwd = 1.2) +
    labs(y = "Relative Growth Rate",
         x = "Root Tissue Density (mg/mm^3)")
#
```
Root-to-Shoot Ratio vs. kwrgr
```{r, echo = T, results='show'}
bsdata_complete %>% 
  ggplot() +
      aes(x = rts, y = kwrgr, color = Origin, shape = Nitrogen) +
      geom_point(na.rm = T) +
      stat_ellipse(na.rm = T, 
                   lwd = 1.2) +
      labs(y = "Relative Growth Rate",
         x = "Root-to-Shoot Ratio")
#Low N has a big effect on rts, but rts has little to no effect on RGR
#most extreme diff in BOYD, COAL has most similar rts and rgr between treatments
```

After D-Lab consultation, 2/9/23
```{r, echo = T, results='show'}
bsdatacat <- bsdata %>% 
  pivot_longer(L01:L10, names_to = 'Lcategory', values_to = 'LengthCat');

#Non-normalized
bsdatacat %>%
    ggplot() +
    aes(x = Lcategory, y = log(LengthCat+1), fill = Origin) +
    geom_boxplot(na.rm = T) +##is this counting zeroes?
    labs(y = "Root Length (log transformed)",
         x = "Root Diameter Class")
#reorder to be by Aridity (BOYD>ELFR>COAL)
# # Box-Cox plot might be helpful!
```
Models, PCA, and MANOVA

PCA (Principal Component Analysis) - Using PRComp
http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/118-principal-component-analysis-in-r-prcomp-vs-princomp/#compute-pca-in-r-using-prcomp
```{r PCA, echo = T, results='show'}
Origin <- bsdata_complete$Origin

#Select relevant root variables
Root_data <- bsdata_complete %>% 
  select(Lengthcm, RootMassg, RootVolcm3, SurfAreacm2, ShootMassg,
         AvgDiammm, rts, srl, sra, kwrgr, rtd)

#cor(Root_data, use = 'complete.obs') #checking each trait's correlation

Root_PCA <- prcomp(na.omit(Root_data), scale = T) 
#scale = T is if they have different units, and na.omit ignores NA values

fviz_eig(Root_PCA)
fviz_pca_biplot(Root_PCA, 
                repel = T, #Avoid text overlapping
                col.ind = "black", # Individuals
                geom = "point",
                alpha.ind = 1, #turns individuals invisible
                addEllipses = T,
                col.var = "#2E9FDF" # Variables color
                ) +
  ylim(-4.5,6.5)

var <- get_pca_var(Root_PCA)
var$contrib

#?fviz_pca_biplot #help
#dim1 is the 'new' set of variables created. Variables are SO highly correlated, PCA collapses them into ONE variable. 
#https://setosa.io/ev/principal-component-analysis/
#note: this looks messy. Could we clean it up?
#from meeting w/ Anusha, 3/10/23 - arrows pointing in same direction are 1.) correlated with each other and 2.) contributing strongly (length of arrow) to dimension 1
#PCA works with colinearity, summarizes what's important into new variables - "What's contributing the most to the variance".
#Look at clustered variables (dim1 = productivity?)
```

LME4 Analysis & Model Building
Length
```{r Model Selection, echo = T, results='show'}
#Building the model
mod_length <- lmer(Lengthcm ~ Origin + Water*Nitrogen + (1 | Parent:Origin), data = bsdata_complete)
#parents form groups within larger groups (Origin)

#summary(mod)
mod_aov <-anova(mod_length)
#mod_aov

# ## Inspect the contrast matrix for the Type III test of Product:
# show_tests(mod_aov)$Origin
# ## Anova-like table of random-effect terms using likelihood ratio tests:
# ranova(mod)
# 
# ## F-tests of 'single term deletions' for all marginal terms:
# drop1(mod)

## backward elimination of non-significant effects:
step_result <- step(mod_length)
# step_result

final_model <- get_model(step_result)
final_model

#Length Model: Lengthcm ~ Origin + (1 | Parent:Origin)

#From my (limited) understanding of the output, only Origin is having a stat signif effect on causing the variation in the model
```

Root-to-Shoot Ratio
```{r, echo = T, results='show'}
mod_rts <- lmer(rts ~ (1|Origin) + Water*Nitrogen + (1 | Sex) + (1 | Parent:Origin), data = bsdata_complete)

summary(mod_rts)
mod_aov <-anova(mod_rts)

step_result <- step(mod_rts)

final_model <- get_model(step_result)
final_model
#formula = rts ~ Origin + Nitrogen
```

RGR
```{r, echo = T, results='show'}
mod_rgr <- lmer(kwrgr ~ (1|Origin) + Water*Nitrogen + (1 | Sex) + (1 | Parent:Origin), data = bsdata_complete)

summary(mod_rgr)
mod_aov <-anova(mod_rgr)

step_result <- step(mod_rgr)

final_model <- get_model(step_result)
final_model
#formula = kwrgr ~ Origin + Water + Nitrogen + (1 | Parent:Origin) + Water:Nitrogen
```
SRL
```{r, echo = T, results='show'}
mod_srl <- lmer(srl ~ Origin + Water*Nitrogen + (1 | Sex) + (1 | Parent:Origin), data = bsdata_complete)

summary(mod_srl)
mod_aov <-anova(mod_srl)

step_result <- step(mod_srl)

final_model <- get_model(step_result)
final_model
#formula = srl ~ 1
```
RTD
```{r, echo = T, results='show'}
mod_rtd <- lmer(rtd ~ AIKoppen + Water*Nitrogen + (1 | Parent:Origin), data = bsdata_complete) #not for this ex., only RDPI
#added in AI Koppen

summary(mod_rtd)
mod_aov <-anova(mod_rtd)

step_result <- step(mod_rtd)

final_model <- get_model(step_result)
final_model
#formula = rtd ~ Water + Nitrogen + (1 | Parent:Origin)
```


```{r MANOVA, echo = T, results ='hide'}
#Approach #1, no random effects 
#Matrix of response variables
Y <- as.matrix(bsdata_complete[, c("rtd", "rts", "srl", "kwrgr",  "sra", "Lengthcm", "RootMassg", "ShootMassg", "RootVolcm3", "SurfAreacm2", "AvgDiammm")])

#Fit MANOVA model with Independent Factors
fit <- manova(Y ~ Nitrogen*Water + AIKoppen + Sex, 
              data = bsdata_complete)
summary(fit)
#Note: N, AIKoppen, and Sex all have statsig p-values

#Approach 2: Adding in Parent random effect 
#Imputing data to estimate NA values
# imputed <- mice(bsdata_complete, m=5, maxit = 50, method = "pmm", predictorMatrix = NULL, seed = 1398)

#Fitting mixed-effects model with random effect of parent
# fitmem <- lmer(cbind("Lengthcm", "rts") ~ 
#                Nitrogen*Water + AIKoppen + Sex + (1|Parent),
#                data = bsdata_complete)
#ASK about imputing data here to deal with NA values

```

For next time: plasticity via RDPI
## https://github.com/ameztegui/Plasticity/blob/master/Plasticity_0.2.pdf

```{r RDPI-RTS, echo = T, results ='hide'}
#from source code at https://rdrr.io/github/ameztegui/Plasticity/src/R/RDPI.R

#From Aitor's email
#data = "bsdata_complete"
#sp = "Origin" (in lieu of species, this is my grouping)
#trait = self-explanatory, here I use Root-to-Shoot ratio "rts"
#factor = "Treatment"

bsdata_complete$Origin <- factor(bsdata_complete$Origin)
bsdata_complete$Treatment <- factor(bsdata_complete$Treatment)
str(bsdata_complete)

#rdpi <- function (data, sp, Trait, factor)
rdpi_rts <- rdpi(bsdata_complete, Origin, rts, Treatment)

#how to deal with missingness of data?



```










